---
title: "CCES Coverage"
author: "Julia Christensen"
date: "April 9, 2018"
output: html_document
---

# PURPOSE OF SCRIPT:
1. Opens cumulative CCES files.  
2. Calculate number of CCES repondents by county.  
3. Plot number of CCES repondents by county.  
4. Repeat #2 and #3 for each state 


# Initial Set Up

- Note: Run CCES_Download_Data first to download and format cces data

```{r setup, include=TRUE}

### Clear global environment
rm(list=ls()) 

### Libraries:
library(pacman)
p_load(tidyverse, foreign, readstata13, readr, haven, foreign, 
       magrittr, mapproj, ggmap, #DeducerSpatial,
       rgdal, RColorBrewer, data.table, tmap, 
       maptools, tmaptools)


### File Locations
file_path_root <- "C:/Users/Julia/OneDrive/Documents/Berkeley/2018_Spring/PS239T_Intro_to_Comp_Tools/ps239T-final-project/"
file_path_proj <- paste0(file_path_root,"Geo_Project_PS239T/")
file_path_dat <- paste0(file_path_root,"Data_Geo_Sophistication/")

### Set theme
my_theme <- theme(
  legend.position = "bottom",
  panel.background = element_rect(fill = NA),
  panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey85"),
  panel.grid.major = element_line(color = "grey95", size = 0.2),
  panel.grid.minor = element_line(color = "grey95", size = 0.2),
  legend.key = element_blank())

```

*****************************************************************

# Load CCES Data

```{r}

### Load RData file with Cumulative CCES dataset
load(file = paste0(file_path_proj,"Data_R/cess_cum.RData"))

```


*****************************************************************

# Plotting CCES Respondents by County

## Download and Clean Map Data

```{r}

### Generate Count Data by County
cces_count <- cces_dat %>% 
  group_by(county_fips) %>%
  summarize(count=n()) %>%
  dplyr::arrange(desc(count)) #%>% print()

### Merge with Map Data
setwd(paste0(file_path_dat,"USDA_Shapefile/"))
US.counties <- readOGR(dsn=".",layer="gz_2010_us_050_00_5m")
US.counties <- US.counties[!(US.counties$STATE %in% c("72")),]  #Remove Puetro Rico
county.data <- US.counties@data
county.data <- cbind(id=rownames(county.data),county.data)
county.data <- data.table(county.data)
county.data[,FIPS:=paste0(STATE,COUNTY)] # this is the state + county FIPS code

setkey(county.data,FIPS)     
cces_count$FIPS <- cces_count$county_fips 
cces_count <- data.table(cces_count)
setkey(cces_count,FIPS)
county.data[cces_count,count:=count]

map.df <- data.table(fortify(US.counties))
setkey(map.df,id)
setkey(county.data,id)
map.df[county.data, count:=count]

setkey(map.df,id)
setkey(county.data,id)
map.df[county.data,  state:=STATE]

### Set Buckets

map.df$count <- ifelse(is.na(map.df$count),0,map.df$count)

map.df$count_scale <- as.factor(cut(map.df$count, 
                                     c(-1, 0, 10, 30, 50, 100, 10000)))

map.df %>% distinct(count_scale)

```

## Generate Map for USA

```{r}

### Subset to mainland USA
#leave out AK, HI, and PR (state FIPS: 02, 15, and 72)
map.df.usa <- map.df[!(map.df$state %in% c("02","15","72")),]

### Save County and State Lines
county_map <- map_data("county")
state_map <- map_data("state")

### Map for USA
ggplot(map.df.usa, aes(x=long, y=lat, group=group)) +
  scale_fill_brewer(palette = "RdYlBu", drop=FALSE, 
                    labels = c("0", "1-10", "11-30", "31-50", "51-100", "100+")) +
  geom_polygon(aes(fill = count_scale)) + 
  geom_path(data = state_map , colour = "white", size=.1) + #Add state lines
  coord_quickmap() +
  labs(title="CCES Responses by County (2006-2016)", x="", y="") +
  theme_void()

### Save map for USA
ggsave(paste0(file_path_proj,"Output/cces_count_USA.pdf"), 
       plot = last_plot(), #Saves last plot
       device = "pdf")

```

*****************************************************************

## Generate Maps for Each U.S. State

```{r}

### Cross Walk for FIPS codes and States
state_fips <- read.table(paste0(file_path_dat,"FIPS_state_codes/state_fips_codes.csv"), 
                         header=TRUE, 
                         sep=",", 
                         colClasses=c("character", "character", "character"))


### Leave out American Samoa, Virgin Islands, Guam, and PR (state FIPS: 60, 78, 66, and 72)
state_fips <- state_fips[!(state_fips$FIPS.Code %in% c("72","60","78","66")),]  
#Note: AK and HI will still not print properly (they will be really small)


### Function for Generateing State Maps
cces_count_state_fun <- function(dat ,state_name, state_code){
  # Create ggplot
  plot <- ggplot(filter(dat, state == as.character(state_code)), #Filter to selected state
       aes(x=long, y=lat, group=group)) +
  scale_fill_brewer(palette = "RdYlBu", drop=FALSE, 
                    labels = c("0", "1-10", "11-30", "31-50", "51-100", "100+")) +
  geom_polygon(aes(fill = count_scale), colour = "white", size=.1) + #Add county lines
  coord_quickmap() +
  labs(title=paste0(state_name," CCES Responses by County (2006-2016)"), x="", y="") +
  theme_void()
  # Return plot
  return(plot)
}


### Generate State maps by looping over each state
for(i in 1:51){
  # Save state names and abbreviations
  name <- state_fips$State.Name[i]
  code <- state_fips$FIPS.Code[i]
  # Generate map 
  ploti <- cces_count_state_fun(dat = map.df, 
                                state_name = name, 
                                state_code = code)
  # Save Map
  ggsave(paste0(file_path_proj,"Output/CCES_Count_States/cces_count_",name,".pdf"), 
         plot = ploti, 
         device = "pdf")
}

```

