---
title: "Mapping USDA Data"
author: "Julia Christensen"
date: "April 9, 2018"
output: html_document
---

# PURPOSE OF SCRIPT:
1. Opens three USDA files.  
2. Plot Rural-Urban Continuum Codes (RUCC) - nonmetropolitan counties by degree of urbanization and adjacency to a metro area (https://www.ers.usda.gov/data-products/rural-urban-continuum-codes/).  
3. Plot Urban Influence Codes (UIC) - distinguishes nonmetropolitan counties by size of the largest city or town and proximity to metro and micropolitan areas (https://www.ers.usda.gov/data-products/urban-influence-codes.aspx).  
4. Plot County Typology  Codes - six mutually exclusive categories of economic dependence and six overlapping categories of policy-relevant themes (https://www.ers.usda.gov/data-products/county-typology-codes.aspx).  


# Note about 1974 Data:

There have been slight changes to county designations since 1974. Before using map generate here in an official capacity, would need to adjust for differences between borders, etc. in 1974 and shape file used here. For now, 1974 data and plots have been commented out.   


# Set Up

```{r setup, include=TRUE}

### Clear global environment
rm(list=ls()) 

### Libraries:
library(pacman)
p_load(tidyverse, foreign, readstata13, readr, haven, foreign, 
       magrittr, mapproj, ggmap, DeducerSpatial,
       rgdal, RColorBrewer, data.table, tmap, 
       maptools, tmaptools, broom, viridis)


### File Locations
file_path_root <- "C:/Users/Julia/OneDrive/Documents/Berkeley/2018_Spring/PS239T_Intro_to_Comp_Tools/ps239T-final-project/"
file_path_proj <- paste0(file_path_root,"Geo_Project_PS239T/")
file_path_dat <- paste0(file_path_root,"Data_Geo_Sophistication/")

### Set theme
my_theme <- theme(
  legend.position = "bottom",
  panel.background = element_rect(fill = NA),
  panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey85"),
  panel.grid.major = element_line(color = "grey95", size = 0.2),
  panel.grid.minor = element_line(color = "grey95", size = 0.2),
  legend.key = element_blank())

```


*****************************************************************

# Load USDA Data

```{r}

### -------------------------------------------------- ###
### File Locations
### -------------------------------------------------- ###

# USDA County Urban Influence Scores 
file_usda_ui <- paste0(file_path_dat,
                       "County_USDA_Urban_Influence/UrbanInfluenceCodes2013.csv")
# USDA County Typology (includes econ type) 
file_usda_tp <- paste0(file_path_dat,
                       "County_USDA_Typology/ERSCountyTypology2015Edition.csv")
# USDA County Rural Urban Continuum Scores
file_usda_ru <- paste0(file_path_dat,
                       "County_USDA_Rural-Urban_Continuum/ruralurbancodes2013.csv")
# # USDA County Rural Urban Continuum Scores from 1974
# file_usda_ru74 <- paste0(file_path_dat,
#                        "County_USDA_Rural-Urban_Continuum/ruralurbancodes1974.csv")

### -------------------------------------------------- ###
### Download & Combine Data
### -------------------------------------------------- ###

### USDA Urban Influence
usda_ui <- read.csv(file = file_usda_ui, 
                      header=TRUE, sep=",") %>% 
  select(-State,-County_Name) %>%
  rename(UIC_Descr = Description)

### USDA Typology
usda_tp <- read.csv(file = file_usda_tp, 
                      header=TRUE, sep=",") #%>% 
  #select(-State,-County_name)

### USDA Rural-Urban
usda_ru <- read.csv(file = file_usda_ru, 
                      header=TRUE, sep=",") %>% 
  select(-State,-County_Name, -Population_2010) %>%
  rename(RUCC_Descr = Description)

# usda_ru74 <- read.csv(file = file_usda_ru74, 
#                       header=TRUE, sep=",") %>% 
#   mutate(FIPS = as.integer(as.character(FIPS.Code))) %>% 
#   rename(RUCC_1974 = X1974.Rural.urban.Continuum.Code) %>%
#   select(FIPS, RUCC_1974) %>%
#   filter(!is.na(FIPS))
  
### Combine USDA Data
usda_all <- full_join(x = usda_ui, 
                      y = usda_tp, 
                      by = "FIPS")
usda_all <- full_join(x = usda_all, 
                      y = usda_ru, 
                      by = "FIPS")
# usda_all <- full_join(x = usda_all,
#                       y = usda_ru74, 
#                       by = "FIPS")

# glimpse(usda_all)

```


*****************************************************************

# Set Up Map File

```{r}

### -------------------------------------------------- ###
### Merge USDA Data with Map Data
### -------------------------------------------------- ###

### Download Map Data
setwd(paste0(file_path_dat,"USDA_Shapefile/"))
US_counties <- readOGR(dsn=".",layer="gz_2010_us_050_00_5m")
#leave out AK, HI, and PR (state FIPS: 02, 15, and 72)
#US.counties <- US.counties[!(US.counties$STATE %in% c("72")),]  #"02","15",
county_data <- US_counties@data
county_data <- cbind(id=rownames(county_data),county_data)
county_data <- as_tibble(county_data)
county_data %<>% mutate(FIPS = as.integer(paste0(STATE,COUNTY)))

### Joing Map data wth USDA Data
county_data <- left_join(x = county_data, 
                         y = usda_all, 
                         by = "FIPS")

map_df <- data.table(broom::tidy(US_counties)) 
map_df <- left_join(x = map_df, 
                         y = county_data, 
                         by = "id")

### Subset to mainland USA
map_df_usa <- map_df[!(map_df$STATE %in% c("02","15","72")),]  

### County and State Borders
map_county <- map_data("county")
map_state <- map_data("state")

```

*****************************************************************

# Maps - UIC

Note for scale_fill_viridis()  
Option - A character string indicating the colormap option to use. Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), and "viridis" (or "D", the default option).  

```{r}

### Define Labels (Unused at the moment)
labels_ui <- c("1"="Metro, Large",
               "2"="Metro, Small",
               "3"="Micro, Adj Large Metro",
               "4"="Noncore, Adj Large Metro",
               "5"="Micro, Adj Small Metro",
               "6"="Noncore (town>2.5k), Adj Large Metro",
               "7"="Noncore (town<2.5k), Adj Large Metro",
               "8"="Micro, Not Adj",
               "9"="Noncore (town>2.5k), Adj Micro",
               "10"="Noncore (town<2.5k), Adj Micro",
               "11"="Noncore (town>2.5k), Adj Noncore",
               "12"="Noncore (town<2.5k), Adj Noncore")

### Plot for 2013 UIC
#plot_ui2013 <- 

plot_ui2013 <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = UIC_2013)) +
  geom_path(data = map_state , colour = "white", size=.1) +
  coord_quickmap() +
  labs(title="USDA Urban Influence Codes by County (2013)", x="", y="") +
  theme_void() 

plot_ui2013_A <- plot_ui2013 + 
  scale_fill_viridis(option="A", direction=1)
plot_ui2013_B <- plot_ui2013 + 
  scale_fill_viridis(option="B", direction=1)
plot_ui2013_D <- plot_ui2013 + 
  scale_fill_viridis(option="D", direction=1)

plot_ui2013_A
plot_ui2013_B
plot_ui2013_D

ggsave(paste0(file_path_proj,"Output/USDA_UIC_2013_USA_A.pdf"), 
       plot = plot_ui2013_A, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_UIC_2013_USA_B.pdf"), 
       plot = plot_ui2013_B, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_UIC_2013_USA_D.pdf"), 
       plot = plot_ui2013_D, device = "pdf")

```

*****************************************************************

# Maps - URCC

```{r}

### Define Labels 
labels_ru <- c("1"="Metro (>1000k)",
               "2"="Metro (250-1000k)",
               "3"="Metro (250-1000k)",
               "4"="Non-Metro (Adj=Y; >20k)",
               "5"="Non-Metro (Adj=N; >20k)",
               "6"="Non-Metro, MAdj (Adj=Y; 2.5-20k)",
               "7"="Non-Metro (Adj=N; 2.5-20k)",
               "8"="Rural (Adj=Y; <2.5k)",
               "9"="Rural (Adj=N; <2.5k)")


### Plot for 2013 URCC
plot_ur2013 <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(RUCC_2013))) +
  geom_path(data = map_state , colour = "white", size=.1) +
  #scale_fill_brewer(palette = "RdYlGn", direction=-1, #YlOrBr
  #                  drop=FALSE, labels = labels_ru) +
  coord_quickmap() +
  labs(title="USDA Rural-Urban Continuum by County (2013)", x="", y="") +
  theme_void() 

plot_ur2013_A <- plot_ur2013 + 
  scale_fill_viridis(option="A", direction=1, discrete=T)
plot_ur2013_B <- plot_ur2013 + 
  scale_fill_viridis(option="B", direction=1, discrete=T)
plot_ur2013_D <- plot_ur2013 + 
  scale_fill_viridis(option="D", direction=1, discrete=T)

plot_ur2013_A
plot_ur2013_B
plot_ur2013_D

ggsave(paste0(file_path_proj,"Output/USDA_RUCC_2013_USA_A.pdf"), 
       plot = plot_ur2013_A, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_RUCC_2013_USA_B.pdf"), 
       plot = plot_ur2013_B, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_RUCC_2013_USA_D.pdf"), 
       plot = plot_ur2013_D, device = "pdf")

# ### Plot for 1974 URCC
# plot_ur1974 <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
#   geom_polygon(aes(fill = as.factor(RUCC_1974))) +
#   geom_path(data = map_state , colour = "white", size=.1) +
#   #scale_fill_brewer(palette = "RdYlGn", direction=-1,
#   #                  drop=FALSE, labels = labels_ru) +
#   coord_quickmap() +
#   labs(title="USDA Rural-Urban Continuum by County (1974)", x="", y="") +
#   theme_void() 
# 
# plot_ur1974_A <- plot_ur1974 + 
#   scale_fill_viridis(option="A", direction=1, discrete=T) 
# plot_ur1974_B <- plot_ur1974 + 
#   scale_fill_viridis(option="B", direction=1, discrete=T) 
# plot_ur1974_D <- plot_ur1974 + 
#   scale_fill_viridis(option="D", direction=1, discrete=T) 
# 
# plot_ur1974_A
# plot_ur1974_B
# plot_ur1974_D
# 
# ggsave(paste0(file_path_proj,"Output/USDA_RUCC_1974_USA_A.pdf"), 
#        plot = plot_ur1974_A, device = "pdf")
# ggsave(paste0(file_path_proj,"Output/USDA_RUCC_1974_USA_B.pdf"), 
#        plot = plot_ur1974_B, device = "pdf")
# ggsave(paste0(file_path_proj,"Output/USDA_RUCC_1974_USA_D.pdf"), 
#        plot = plot_ur1974_D, device = "pdf")

```


*****************************************************************

# Maps - Econ Type

```{r}

### Turn Dummies into Factor
map_df_usa %<>% mutate(Econ_Type="Other") %>%
  mutate(Econ_Type=ifelse(map_df_usa$Farming==1,"Farming",Econ_Type)) %>%
  mutate(Econ_Type=ifelse(map_df_usa$Mining==1,"Mining",Econ_Type)) %>%
  mutate(Econ_Type=ifelse(map_df_usa$Manufacturing==1,"Manufacturing",Econ_Type)) %>%
  mutate(Econ_Type=ifelse(map_df_usa$Government==1,"Government",Econ_Type)) %>%
  mutate(Econ_Type=ifelse(map_df_usa$Recreation==1,"Recreation",Econ_Type)) %>%
  mutate(Econ_Type=ifelse(map_df_usa$Nonspecialized==1,"Nonspecialized",Econ_Type)) 

### Plot for 2013 Econ Type
plot_tp2013 <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = Econ_Type)) +
  geom_path(data = map_state , colour = "white", size=.1) +
  #scale_fill_brewer(palette = "Set1", direction=-1,
  #                  drop=FALSE) +
  coord_quickmap() +
  labs(title="USDA Econ Type by County (2013)", x="", y="") +
  theme_void() 

plot_tp2013 <- plot_tp2013 + 
  scale_fill_manual(values=c("Farming"="#ffd92f",
                             "Mining"="#6a3d9a",
                             "Manufacturing"="#e31a1c",
                             "Government"="#cab2d6",
                             "Recreation"="#33a02c",
                             "Nonspecialized"="#f0f0f0"))

plot_tp2013_alt1 <- plot_tp2013 + 
  scale_fill_brewer(palette = "Set1", direction=-1, drop=FALSE)

plot_tp2013_alt2 <- plot_tp2013 + 
  scale_fill_brewer(palette = "Set3", direction=-1, drop=FALSE)

ggsave(paste0(file_path_proj,"Output/USDA_EconType_2013_USA.pdf"), 
       plot = plot_tp2013, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_EconType_Alt1_2013_USA.pdf"), 
       plot = plot_tp2013_alt1, device = "pdf")
ggsave(paste0(file_path_proj,"Output/USDA_EconType_Alt2_2013_USA.pdf"), 
       plot = plot_tp2013_alt2, device = "pdf")

```


*****************************************************************

# Maps - Other Econ Fields

```{r}

### Plot for 2013 Low_Edu
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Low_Edu))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Low Education by County (2013)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Low_Edu_2013_USA.pdf"), 
       plot = last_plot(), device = "pdf")

### Plot for 2013 Low_Employment
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Low_Employment))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Low Employment by County (2013)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Low_Employment_2013_USA.pdf"), 
       plot = last_plot(), device = "pdf")

### Plot for 2013 Retirement_Dest
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Retirement_Dest))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Retirement Destination by County (2013)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Retirement_Dest_2013_USA.pdf"), 
       plot = last_plot(), device = "pdf")

### Plot for 2013 Pop_Loss_2010
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Pop_Loss_2010))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Population Loss by County (2010)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Pop_Loss_2010_USA.pdf"), 
       plot = last_plot(), device = "pdf")

### Plot for 2013 Persistant_Poverty_2013
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Persistant_Poverty_2013))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Persistant Poverty by County (2013)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Persistant_Poverty_2013_USA.pdf"), 
       plot = last_plot(), device = "pdf")

### Plot for 2013 Persistent_Child_Poverty_2013
ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = as.factor(Persistent_Child_Poverty_2013))) +
  geom_path(data = map_state , colour = "grey", size=.1) +
  scale_fill_manual(values=c("1"="#525252","0"="#f7f7f7")) +
  coord_quickmap() +
  labs(title="USDA Persistant Child Poverty by County (2013)", x="", y="") +
  theme_void() 

ggsave(paste0(file_path_proj,"Output/USDA_Persistant_Child_Poverty_2013_USA.pdf"), 
       plot = last_plot(), device = "pdf")

```