---
title: "Mapping Census Data"
author: "Julia Christensen"
date: "April 23, 2018"
output: html_document
---

# PURPOSE OF SCRIPT:
1. Opens three USDA files.  
- Plot Rural-Urban Continuum Codes (RUCC) - nonmetropolitan counties by degree of urbanization and adjacency to a metro area (https://www.ers.usda.gov/data-products/rural-urban-continuum-codes/).  
- Plot Urban Influence Codes (UIC) - distinguishes nonmetropolitan counties by size of the largest city or town and proximity to metro and micropolitan areas (https://www.ers.usda.gov/data-products/urban-influence-codes.aspx).  
- Plot County Typology  Codes - six mutually exclusive categories of economic dependence and six overlapping categories of policy-relevant themes (https://www.ers.usda.gov/data-products/county-typology-codes.aspx).  
2. Download Census data.  
3. Merge USDA and Census data.  
4. Plot population density.  


# Set Up

```{r setup, include=TRUE}

knitr::opts_chunk$set(echo = TRUE)

### Clear global environment
rm(list=ls()) 

### Libraries:
library(pacman)
p_load(tidyverse, foreign, readstata13, readr, haven, foreign, 
       magrittr, mapproj, ggmap, ggthemes, #DeducerSpatial,
       rgdal, RColorBrewer, data.table, tmap, 
       #maptools, tmaptools, 
       broom, viridis, 
       tidycensus, census)


### Store API Key (for tidycensus package)
# Get API Key here: https://api.census.gov/data/key_signup.html
jbc_census_api_key <- "[INSERT KEY HERE]"
tidycensus::census_api_key(key = jbc_census_api_key, 
                           overwrite = FALSE, 
                           install = FALSE)

### File Locations
file_path_root <- "C:/Users/Julia/OneDrive/Documents/Berkeley/2018_Spring/PS239T_Intro_to_Comp_Tools/ps239T-final-project/"
file_path_proj <- paste0(file_path_root,"Geo_Project_PS239T/")
file_path_dat <- paste0(file_path_root,"Data_Geo_Sophistication/")

### Set theme
my_theme <- theme(
  legend.position = "bottom",
  panel.background = element_rect(fill = NA),
  panel.border = element_rect(fill = NA, color = "grey75"),
  axis.ticks = element_line(color = "grey85"),
  panel.grid.major = element_line(color = "grey95", size = 0.2),
  panel.grid.minor = element_line(color = "grey95", size = 0.2),
  legend.key = element_blank())

```


*****************************************************************

# Load USDA Data on County Features

```{r}

### -------------------------------------------------- ###
### File Locations
### -------------------------------------------------- ###

# USDA County Urban Influence Scores 
file_usda_ui <- paste0(file_path_dat,
                       "County_USDA_Urban_Influence/UrbanInfluenceCodes2013.csv")
# USDA County Typology (includes econ type) 
file_usda_tp <- paste0(file_path_dat,
                       "County_USDA_Typology/ERSCountyTypology2015Edition.csv")
# USDA County Rural Urban Continuum Scores
file_usda_ru <- paste0(file_path_dat,
                       "County_USDA_Rural-Urban_Continuum/ruralurbancodes2013.csv")
# USDA County Rural Urban Continuum Scores from 1974
file_usda_ru74 <- paste0(file_path_dat,
                       "County_USDA_Rural-Urban_Continuum/ruralurbancodes1974.csv")

### -------------------------------------------------- ###
### Download & Combine Data
### -------------------------------------------------- ###

### USDA Urban Influence
usda_ui <- read.csv(file = file_usda_ui, 
                      header=TRUE, sep=",") %>% 
  select(-State,-County_Name) %>%
  rename(UIC_Descr = Description)

### USDA Typology
usda_tp <- read.csv(file = file_usda_tp, 
                      header=TRUE, sep=",") #%>% 
  #select(-State,-County_name)

### USDA Rural-Urban
usda_ru <- read.csv(file = file_usda_ru, 
                      header=TRUE, sep=",") %>% 
  select(-State,-County_Name, -Population_2010) %>%
  rename(RUCC_Descr = Description)

usda_ru74 <- read.csv(file = file_usda_ru74, 
                      header=TRUE, sep=",") %>% 
  mutate(FIPS = as.integer(as.character(FIPS.Code))) %>% 
  rename(RUCC_1974 = X1974.Rural.urban.Continuum.Code) %>%
  select(FIPS, RUCC_1974) %>%
  filter(!is.na(FIPS))
  
### Combine USDA Data
usda_all <- full_join(x = usda_ui, 
                      y = usda_tp, 
                      by = "FIPS")
usda_all <- full_join(x = usda_all, 
                      y = usda_ru, 
                      by = "FIPS")
usda_all <- full_join(x = usda_all,
                      y = usda_ru74, 
                      by = "FIPS")

# glimpse(usda_all)
```

*****************************************************************

# Load Census Population Data

## Using CSV from Census Website
```{r}

### Census Rural Urban Pop Totals 
# Source: https://www.census.gov/geo/reference/urban-rural.html

# File Path 
file_census_ru <- paste0(file_path_dat,
                         "County_Rural_Census/County_Rural_Lookup.csv")

# Download data
census_ru <- read.csv(file = file_census_ru, 
                      header=TRUE, 
                      sep=",",
                      colClasses = c("character",
                                     "character",
                                     "character",
                                     "numeric",
                                     "numeric",
                                     "numeric",
                                     "numeric")) %>% 
  mutate(FIPS = as.integer(fips_code)) %>% 
  mutate(state_code = as.integer(substr(fips_code, start = 1, stop = 2))) %>%
  print()

### Add summary by state
census_ru_state_totals <- census_ru %>% 
  group_by(state_code) %>%
  summarise(state_total_pop = sum(total_pop),
            state_total_pop_urban = sum(total_pop_urban),
            state_total_pop_rural = sum(total_pop_rural),
            state_percent_rural = sum(total_pop_rural)/sum(total_pop) ) %>%
  ungroup() %>%
  print()

census_ru <- left_join(x = census_ru, 
                       y = census_ru_state_totals, 
                       by = "state_code")

```

## Using Tidycensus R package and census API

Sources:  
https://walkerke.github.io/tidycensus/articles/spatial-data.html  
https://www.pdx.edu/prc/sites/www.pdx.edu.prc/files/06_pragmatic_tidy.pdf  
http://knaaptime.com/projects/neighborhood-types/opportunity-types-in-r/  
http://dlab.berkeley.edu/blog/season-sharing-data-working-newly-released-census-2010-2014-acs-5-year-data-r  
https://www.gl-li.com/  
https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml  
http://edrub.in/ARE212/section12.html#spatial_data  
https://edzer.github.io/UseR2017/  

```{r}

# ### ---
# ### Note:
# ### The code below should download the data summarized in the CSV above, but for some reason
# ### the census API is returning zeros for all the values other than total population/housing units
# ### ---
# 
# 
# ### Save List of Variables from ACS and Census
# # Use View() and filter within view to search variables 
# vars_acs5_16 <- tidycensus::load_variables(year = 2016, #year of Census or endyear of ACS
#                                   dataset = "acs5", #options: "sf1", "sf3", or "acs5"
#                                   cache = TRUE)
# 
# vars_sf1_10 <- tidycensus::load_variables(year = 2010, #year of Census or endyear of ACS
#                                   dataset = "sf1", #options: "sf1", "sf3", or "acs5"
#                                   cache = TRUE)
# # See also http://wagda.lib.washington.edu/data/type/census/geodb/metadata/SF1qkRef_2010.pdf
# 
# 
# ### Download 2010 Housing Units Data from Census
# census_housing_vars <- c(housing_units = "H00010001",
#                          housing_units2 = "H0020001",
#                          housing_units_urban = "H0020002",
#                          housing_units_urban_area = "H0020003", #Urban: !! Inside urbanized areas
#                          housing_units_urban_cluster_u = "H0020004", #Urban: !! Inside urban clusters
#                          housing_units_urban_cluster_r = "H0020005") #Rural !! Inside urban clusters
# census_housing_vars_sum <- c("H00010001")
# 
# 
# housing_2010 <- get_decennial(geography = "county", 
#                               variables = census_housing_vars,
#                               year = 2010,
#                               geometry = FALSE,
#                               summary_var = census_housing_vars_sum) %>%
#   mutate(FIPS = as.integer(GEOID))
# 
# housing_2010_wide <- housing_2010 %>% 
#   spread(key = variable,
#          value = value)
# 
# 
# ### Download 2010 Population Data from Census
# census_pop_vars <- c(pop = "P0010001",
#                      pop2 = "P0020001",
#                      pop_urban = "P0020002",
#                      pop_urban_area = "P0020003", #Urban: !! Inside urbanized areas
#                      pop_urban_cluster_u = "P0020004", #Urban: !! Inside urban clusters
#                      pop_urban_cluster_r = "P0020005") #Rural !! Inside urban clusters
# census_pop_vars_sum <- c("P0010001")
# 
# pop_2010 <- get_decennial(geography = "county", 
#                           variables = census_pop_vars,
#                           year = 2010,
#                           geometry = FALSE,
#                           summary_var = census_pop_vars_sum) %>%
#   mutate(FIPS = as.integer(GEOID))
# 
# pop_2010_wide <- pop_2010 %>% 
#   spread(key = variable,
#          value = value)
# 
# sum(pop_2010_wide$pop_urban_area)

```



*****************************************************************

# Combine Data

```{r}

### -------------------------------------------------- ###
### Merge USDA and Census Data with Map Data
### -------------------------------------------------- ###

### Download Map Data
setwd(paste0(file_path_dat,"USDA_Shapefile/"))
US_counties <- readOGR(dsn=".",layer="gz_2010_us_050_00_5m")
#leave out AK, HI, and PR (state FIPS: 02, 15, and 72)
#US.counties <- US.counties[!(US.counties$STATE %in% c("72")),]  #"02","15",
county_data <- US_counties@data
county_data <- cbind(id=rownames(county_data),county_data)
county_data <- as_tibble(county_data)
county_data %<>% mutate(FIPS = as.integer(paste0(STATE,COUNTY)))

### Joining County data wth USDA Data
county_data <- left_join(x = county_data, 
                         y = usda_all, 
                         by = "FIPS")

### Joining County data wth Census Data
county_data <- left_join(x = county_data, 
                         y = census_ru, 
                         by = "FIPS")

### Create map dataframe
map_df <- data.table(broom::tidy(US_counties)) 
map_df <- left_join(x = map_df, 
                    y = county_data, 
                    by = "id")

### Subset to mainland USA
map_df_usa <- map_df[!(map_df$STATE %in% c("02","15","72")),]  

### County and State Borders
map_county <- map_data("county")
map_state <- map_data("state")

```


*****************************************************************

# Maps - Census by State

```{r}

### Create base plot by county
plot_pct_rural_2010_s <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = state_percent_rural)) +
  geom_path(data = map_state , colour = "white", size=.1) +
  coord_quickmap() +
  labs(title="% Rural Population by State, 2010 Census", x="", y="") +
  theme_void() 

### Create plots with different color scales
plot_pct_rural_2010_s_A <- plot_pct_rural_2010_s + 
  scale_fill_viridis(option="A", direction=1)
plot_pct_rural_2010_s_B <- plot_pct_rural_2010_s + 
  scale_fill_viridis(option="B", direction=1)
plot_pct_rural_2010_s_D <- plot_pct_rural_2010_s + 
  scale_fill_viridis(option="D", direction=1)

### View plots with different color scales
plot_pct_rural_2010_s_A
plot_pct_rural_2010_s_B
plot_pct_rural_2010_s_D

### Save plots with different color scales
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_s_USA_A.pdf"), 
       plot = plot_pct_rural_2010_s_A, device = "pdf")
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_s_USA_B.pdf"), 
       plot = plot_pct_rural_2010_s_B, device = "pdf")
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_s_USA_D.pdf"), 
       plot = plot_pct_rural_2010_s_D, device = "pdf")

```


*****************************************************************

# Maps - Census by County

```{r}

### Create base plot by county
plot_pct_rural_2010 <- ggplot(map_df_usa, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill = percent_rural)) +
  geom_path(data = map_state , colour = "white", size=.1) +
  coord_quickmap() +
  labs(title="% Rural Population by County, 2010 Census", x="", y="") +
  theme_void() 

### Create plots with different color scales
plot_pct_rural_2010_A <- plot_pct_rural_2010 + 
  scale_fill_viridis(option="A", direction=1)
plot_pct_rural_2010_B <- plot_pct_rural_2010 + 
  scale_fill_viridis(option="B", direction=1)
plot_pct_rural_2010_D <- plot_pct_rural_2010 + 
  scale_fill_viridis(option="D", direction=1)

### View plots with different color scales
plot_pct_rural_2010_A
plot_pct_rural_2010_B
plot_pct_rural_2010_D

### Save plots with different color scales
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_USA_A.pdf"), 
       plot = plot_pct_rural_2010_A, device = "pdf")
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_USA_B.pdf"), 
       plot = plot_pct_rural_2010_B, device = "pdf")
ggsave(paste0(file_path_proj,"Output/Census_rural_2010_USA_D.pdf"), 
       plot = plot_pct_rural_2010_D, device = "pdf")

```


*****************************************************************

# Maps - USDA Est. by State

```{r}

### -------------------------------------------------- ###
### Create Estimates using URCC 
### -------------------------------------------------- ###

### Recall RUCC_2013 Labels 
labels_ru <- c("1"="Metro (>1000k)",
               "2"="Metro (250-1000k)",
               "3"="Metro (250-1000k)",
               "4"="Non-Metro (Adj=Y; >20k)",
               "5"="Non-Metro (Adj=N; >20k)",
               "6"="Non-Metro, MAdj (Adj=Y; 2.5-20k)",
               "7"="Non-Metro (Adj=N; 2.5-20k)",
               "8"="Rural (Adj=Y; <2.5k)",
               "9"="Rural (Adj=N; <2.5k)")

### Create estimate of rural pop based on USDA RUCC
map_df %<>% 
  mutate(rural_RUCC_4 = ifelse(RUCC_2013 > 3, 1, 0) ) %>%
  mutate(est_rural_pop_4 = rural_RUCC_4*total_pop) %>%
  
  mutate(rural_RUCC_5 = ifelse(RUCC_2013 > 4, 1, 0) ) %>%
  mutate(est_rural_pop_5 = rural_RUCC_5*total_pop) %>%
  
  mutate(rural_RUCC_6 = ifelse(RUCC_2013 > 5, 1, 0) ) %>%
  mutate(est_rural_pop_6 = rural_RUCC_6*total_pop) %>%
  
  mutate(rural_RUCC_7 = ifelse(RUCC_2013 > 6, 1, 0) ) %>%
  mutate(est_rural_pop_7 = rural_RUCC_7*total_pop)
  
### Add summary by state
map_df_state_totals <- map_df %>% 
  group_by(state_code) %>%
  summarise(state_percent_rural_URCC_est_4 = sum(est_rural_pop_4)/sum(total_pop),
            state_percent_rural_URCC_est_5 = sum(est_rural_pop_5)/sum(total_pop),
            state_percent_rural_URCC_est_6 = sum(est_rural_pop_6)/sum(total_pop),
            state_percent_rural_URCC_est_7 = sum(est_rural_pop_7)/sum(total_pop)) 

map_df<- left_join(x = map_df, 
                       y = map_df_state_totals, 
                       by = "state_code")

### Subset to continential states
map_df_usa <- map_df[!(map_df$STATE %in% c("02","15","72")),] 


### -------------------------------------------------- ###
### Create Estimates using UIC 
### -------------------------------------------------- ###

### Recall UIC_2013 Labels 
labels_ui <- c("1"="Metro, Large",
               "2"="Metro, Small",
               "3"="Micro, Adj Large Metro",
               "4"="Noncore, Adj Large Metro",
               "5"="Micro, Adj Small Metro",
               "6"="Noncore (town>2.5k), Adj Large Metro",
               "7"="Noncore (town<2.5k), Adj Large Metro",
               "8"="Micro, Not Adj",
               "9"="Noncore (town>2.5k), Adj Micro",
               "10"="Noncore (town<2.5k), Adj Micro",
               "11"="Noncore (town>2.5k), Adj Noncore",
               "12"="Noncore (town<2.5k), Adj Noncore")

### Create estimate of rural pop based on USDA RUCC
map_df %<>% 
  mutate(rural_UICC_6 = ifelse(UIC_2013 > 5, 1, 0) ) %>%
  mutate(est_rural_pop_UICC_6 = rural_UICC_6*total_pop) %>%
  
  mutate(rural_UICC_7 = ifelse(UIC_2013 > 6, 1, 0) ) %>%
  mutate(est_rural_pop_UICC_7 = rural_UICC_7*total_pop) %>%
  
  mutate(rural_UICC_8 = ifelse(UIC_2013 > 7, 1, 0) ) %>%
  mutate(est_rural_pop_UICC_8 = rural_UICC_8*total_pop) %>%
  
  mutate(rural_UICC_9 = ifelse(UIC_2013 > 8, 1, 0) ) %>%
  mutate(est_rural_pop_UICC_9 = rural_UICC_9*total_pop)
   
  
### Add summary by state
map_df_state_totals <- map_df %>% 
  group_by(state_code) %>%
  summarise(state_percent_rural_UICC_est_6 = sum(est_rural_pop_UICC_6)/sum(total_pop),
            state_percent_rural_UICC_est_7 = sum(est_rural_pop_UICC_7)/sum(total_pop),
            state_percent_rural_UICC_est_8 = sum(est_rural_pop_UICC_8)/sum(total_pop),
            state_percent_rural_UICC_est_9 = sum(est_rural_pop_UICC_9)/sum(total_pop)) 

map_df<- left_join(x = map_df, 
                       y = map_df_state_totals, 
                       by = "state_code")

### Subset to continential states
map_df_usa <- map_df[!(map_df$STATE %in% c("02","15","72")),] 


### -------------------------------------------------- ###
### Map using URCC and UIC 
### -------------------------------------------------- ###


### Create function for plotting by state
plot_USDA <- function(number, scale, dat, fill_var){
  plot1 <- ggplot(dat, aes(x=long, y=lat, group=group)) +
    geom_polygon(aes_string(fill = fill_var)) +
    geom_path(data = map_state , colour = "white", size=.1) +
    coord_quickmap() +
    labs(title=paste0("% Rural Population, USDA ",scale," > ",number-1), 
         x="", y="") +
    theme_void() + 
    scale_fill_viridis(option="B", direction=1)
  
  ggsave(paste0(file_path_proj,"Output/USDA_",scale,number,"_rural_2010_USA_B.pdf"), 
       plot = plot1, device = "pdf")
  
  return(plot1)
}

### URCC Maps
plot_USDA(number = 4,
          scale = "URCC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_URCC_est_4")

plot_USDA(number = 5,
          scale = "URCC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_URCC_est_5")

plot_USDA(number = 6,
          scale = "URCC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_URCC_est_6")

plot_USDA(number = 7,
          scale = "URCC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_URCC_est_7")


### UIC Maps
plot_USDA(number = 6,
          scale = "UIC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_UICC_est_6")

plot_USDA(number = 7,
          scale = "UIC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_UICC_est_7")

plot_USDA(number = 8,
          scale = "UIC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_UICC_est_8")

plot_USDA(number = 9,
          scale = "UIC", 
          dat = map_df_usa, 
          fill_var = "state_percent_rural_UICC_est_9")

```

```{r}

### Comparing USDA scales to Census rural designation

# UIC vs. Census (plot then save)
ggplot(county_data, aes(x=UIC_2013, y=percent_rural)) + 
  geom_point(shape = 21, alpha=.5) +
  geom_smooth(method = "lm",
              span = .1,  # Add fitted line 
              formula = y ~ x, #Source: http://ggplot2.tidyverse.org/reference/geom_smooth.html
              se = TRUE, 
              color="red") + 
  ggtitle("All Counties, UIC (USDA) vs. %Rural (Census)") +
  my_theme

ggsave(paste0(file_path_proj,"Output/Compare_UIC_v_Census_Rural.pdf"), 
       plot = last_plot(), device = "pdf")

# RUCC vs. Census (plot then save)
ggplot(county_data, aes(x=RUCC_2013, y=percent_rural)) + 
  geom_point(shape = 21, alpha=.5) +
  geom_smooth(method = "lm",
              span = .1,  # Add fitted line 
              formula = y ~ x, #Source: http://ggplot2.tidyverse.org/reference/geom_smooth.html
              se = TRUE, 
              color="red") + 
  ggtitle("All Counties, RUCC (USDA) vs. %Rural (Census)") +
  my_theme

ggsave(paste0(file_path_proj,"Output/Compare_RUCC_v_Census_Rural.pdf"), 
       plot = last_plot(), device = "pdf")



### What's going on with Colorado?
# Colorado is the most noticable difference between Census and USDA estimates in the above plots 
co_data <- county_data %>%
  select(abb, UIC_2013, RUCC_2013, total_pop, percent_rural) %>%
  filter(abb=="CO")

ggplot(co_data, aes(x=UIC_2013, y=percent_rural)) + 
  geom_point(shape = 21, alpha=.5) +
  geom_smooth(method = "lm",
              span = .1,  # Add fitted line 
              formula = y ~ x, #Source: http://ggplot2.tidyverse.org/reference/geom_smooth.html
              se = TRUE, 
              color="red") + 
  ggtitle("CO Counties, UIC vs. %Rural") +
  my_theme

ggplot(co_data, aes(x=RUCC_2013, y=percent_rural)) + 
  geom_point(shape = 21, alpha=.5) +
  geom_smooth(method = "lm",
              span = .1,  # Add fitted line 
              formula = y ~ x, #Source: http://ggplot2.tidyverse.org/reference/geom_smooth.html
              se = TRUE, 
              color="red") + 
  ggtitle("CO Counties, URCC vs. %Rural") +
  my_theme

```


